---
title: "Dashboard"
format: 
  html:
    page-layout: full
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
library(haven)
library(dplyr)
library(ggplot2)
library(knitr)
library(scales)
library(DT)

demo_data <- read_xpt("DEMO_L.XPT")

demo_cleaned <- demo_data %>%
  mutate(
    income_4cat = case_when(
      INDFMPIR < 1.30                     ~ "Low Income",
      INDFMPIR >= 1.30 & INDFMPIR < 3.50  ~ "Middle-Low Income",
      INDFMPIR >= 3.50 & INDFMPIR < 5.00  ~ "Middle-High Income",
      INDFMPIR >= 5.00                    ~ "High Income",
      TRUE                                ~ NA_character_
    ) %>% factor(
      levels = c("Low Income", "Middle-Low Income", "Middle-High Income", "High Income")
    ),
    gender = factor(
      ifelse(RIAGENDR == 1, "Male", "Female"),
      levels = c("Male", "Female")
    )
  ) %>%
  filter(!is.na(income_4cat), !is.na(gender))

navy_blue     <- "#12263F"
golden_yellow <- "#B88E10"

total_participants <- nrow(demo_cleaned)
low_income_pct     <- mean(demo_cleaned$income_4cat == "Low Income") * 100
high_income_pct    <- mean(demo_cleaned$income_4cat == "High Income") * 100

income_summary <- demo_cleaned %>%
  count(income_4cat) %>%
  mutate(proportion = n / sum(n))

data_table <- demo_cleaned %>%
  group_by(`Income Category` = income_4cat) %>%
  summarise(
    `Total Participants` = n(),
    `Mean Age` = round(mean(RIDAGEYR, na.rm = TRUE), 1),
    `Proportion (%)` = sprintf("%.1f%%", n() / nrow(demo_cleaned) * 100)
  )
```

## {.sidebar}

This dashboard displays statistics for:

**Dataset**\
NHANES 2021-2023

**File**\
DEMO_L.XPT

**Component**\
Demographics

------------------------------------------------------------------

**Total records**\
`r nrow(demo_data)`

**Records used (after filtering NA)**\
`r nrow(demo_cleaned)`

------------------------------------------------------------------

::: {.callout-note collapse="true"}
## Disclaimer
Income categories are derived from the family income-to-poverty ratio (`INDFMPIR`). Records with missing income or gender values are excluded.
:::

# ALL

## Row {height="22%"}

```{r}
#| content: valuebox
#| title: "Total participants"
#| icon: people
#| color: "#7C93AC"
list(value = total_participants)
```

```{r}
#| content: valuebox
#| title: "Low income"
#| icon: graph-down
#| color: "#B36B6B"
list(value = paste0(round(low_income_pct, 1), "%"))
```

```{r}
#| content: valuebox
#| title: "High income"
#| icon: graph-up
#| color: "#B88E10"
list(value = paste0(round(high_income_pct, 1), "%"))
```

## Row {height="40%"}

### Column {width="45%"}

#### Income summary

```{r}
data_table %>%
  kable(align = c("l", "r", "r", "r"))
```

### Column {width="55%"}

#### Family Income Category Distribution

```{r}
ggplot(income_summary, aes(x = income_4cat, y = proportion, fill = income_4cat)) +
  geom_col(width = 0.6, show.legend = FALSE) +
  geom_text(
    aes(label = paste0(n, " (", percent(proportion, accuracy = 0.1), ")")),
    vjust = -0.5,
    size = 3.5,
    fontface = "bold"
  ) +
  scale_y_continuous(labels = percent_format(), limits = c(0, max(income_summary$proportion) * 1.15)) +
  scale_fill_brewer(palette = "Set2") +
  labs(x = NULL, y = "Proportion") +
  theme_minimal(base_size = 12) +
  theme(panel.grid.major.x = element_blank())
```

## Row {height="38%"}

#### Gender Composition Across Income Tiers

```{r}
ggplot(demo_cleaned, aes(y = income_4cat, fill = gender)) +
  geom_bar(position = "fill", width = 0.6) +
  scale_x_continuous(labels = percent_format(), name = "Proportion") +
  scale_y_discrete(limits = rev(levels(demo_cleaned$income_4cat)), name = NULL) +
  scale_fill_manual(values = c("Male" = navy_blue, "Female" = golden_yellow), name = "Gender") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "right", panel.grid.major.y = element_blank())
```

# DATA

## Row

```{r}
datatable(
  demo_cleaned,
  options = list(pageLength = 10, scrollX = TRUE, autoWidth = TRUE),
  caption = "Full cleaned NHANES 2021-2023 demographic dataset",
  rownames = FALSE
)
```
